---
title: "Data analysis"
author: "JasonL"
date: "31/03/2022"
output: github_document
---

SECTION 1: Load packages we need.
```{r Packages, message=FALSE}
library(reshape2)
library(dplyr)
library(plyr)
library(ggplot2)
library(lme4)
library(RLRsim)
library(car)
```


SECTION 2: Upload the data, name it, and look at it.
```{r Data}
setwd("C:/Users/jason/Dropbox/My PC (DESKTOP-IOO274E)/Desktop/Data/DuckweedExpEvol/Git_repo/Lemna_exp_evol")

data<-read.table('DW_Exp_Evol_final.txt', na.strings="", header=FALSE, sep='\t')

#There are alot of columns to label here. 
names(data)<-c('Pop','Media','Trt','Rep','G1ID','G1Pos','G1Edge','G1_notes0','G1_fr0','G1_pix0','G1_per0','G1_grn0','G1_grnper0','G1_notes1','G1_fr1','G1_pix1','G1_per1','G1_grn1','G1_grnper1','G1_abscnt','G2ID','G2Pos','G2Edge','G2_notes0','G2_fr0','G2_pix0','G2_per0','G2_grn0','G2_grnper0','G2_notes1','G2_fr1','G2_pix1','G2_per1','G2_grn1','G2_grnper1','G2_abscnt','G3ID','G3Pos','G3Edge','G3_notes0','G3_fr0','G3_pix0','G3_per0','G3_grn0','G3_grnper0','G3_notes1','G3_fr1','G3_pix1','G3_per1','G3_grn1','G3_grnper1','G3_abscnt','G4ID','G4Pos','G4Edge','G4_notes0','G4_fr0','G4_pix0','G4_per0','G4_grn0','G4_grnper0','G4_notes1','G4_fr1','G4_pix1','G4_per1','G4_grn1','G4_grnper1','G4_abscnt','G5ID','G5Pos','G5Edge','G5_notes0','G5_fr0','G5_pix0','G5_per0','G5_grn0','G5_grnper0','G5_notes1','G5_fr1','G5_pix1','G5_per1','G5_grn1','G5_grnper1','G5_abscnt','G6ID','G6Pos','G6Edge','G6_notes0','G6_fr0','G6_pix0','G6_per0','G6_grn0','G6_grnper0','G6_notes1','G6_fr1','G6_pix1','G6_per1','G6_grn1','G6_grnper1','G6_abscnt','G7ID','G7Pos','G7Edge','G7_notes0','G7_fr0','G7_pix0','G7_per0','G7_grn0','G7_grnper0','G7_notes1','G7_fr1','G7_pix1','G7_per1','G7_grn1','G7_grnper1','G7_abscnt','G8ID','G8Pos','G8Edge','G8_notes0','G8_fr0','G8_pix0','G8_per0','G8_grn0','G8_grnper0','G8_notes1','G8_fr1','G8_pix1','G8_per1','G8_grn1','G8_grnper1','G8_abscnt','G9ID','G9Pos','G9Edge','G9_notes0','G9_fr0','G9_pix0','G9_per0','G9_grn0','G9_grnper0','G9_notes1','G9_fr1','G9_pix1','G9_per1','G9_grn1','G9_grnper1','G9_abscnt', 'G10ID','G10Pos','G10Edge','G10_notes0','G10_fr0','G10_pix0','G10_per0','G10_grn0','G10_grnper0','G10_notes1','G10_fr1','G10_pix1','G10_per1','G10_grn1','G10_grnper1','G10_abscnt')

head(data)
```

SECTION 3: Data processing.
Let's clean some of this up. This is where figure legends can be adjusted, etc. 
```{r data processing}
#Generate legend labels for treatment and populations.
data <- data.frame(lapply(data, function(x) {
  gsub("RegularK", "Ctl", x)
}))
data <- data.frame(lapply(data, function(x) {
  gsub("SaltK", "Salt", x)
}))
data <- data.frame(lapply(data, function(x) {
  gsub("PLT", "Plt", x)
}))
data <- data.frame(lapply(data, function(x) {
  gsub("BAC", "Bac", x)
})) 
data <- data.frame(lapply(data, function(x) {
  gsub("Bac_Plt", "Holo", x)
})) 
data <- data.frame(lapply(data, function(x) {
  gsub("Wellspring", "W", x)
})) 
data <- data.frame(lapply(data, function(x) {
  gsub("Churchill", "C", x)
}))
data <- data.frame(lapply(data, function(x) {
  gsub("RougePark", "R", x)
})) 

# Break dataset into different generations so that we can realign the data
Gen1<- data[,1:20]
Gen2<- data[,c(1:5,22:36)]
Gen3<- data[,c(1:5,38:52)]
Gen4<- data[,c(1:5,54:68)]
Gen5<- data[,c(1:5,70:84)]
Gen6<- data[,c(1:5,86:100)]
Gen7<- data[,c(1:5,102:116)]
Gen8<- data[,c(1:5,118:132)]
Gen9<- data[,c(1:5,134:148)]
Gen10<- data[,c(1:5,150:164)]

# Reassign variable names.
names(Gen1)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen2)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen3)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen4)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen5)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen6)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen7)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen8)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen9)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')
names(Gen10)<-c('Pop','Med','Trt','Rep','g1ID','Pos','Edge','Notes0','Fr0','Pix0','Per0','Grn0','GrnPer0','Notes1','Fr1','Pix1','Per1','Grn1','GrnPer1','AbsCnt')

# Assemble into a single dataset
CWR_1to10<-rbind(Gen1,Gen2,Gen3,Gen4,Gen5,Gen6,Gen7,Gen8,Gen9,Gen10)
CWR_1to10$Gen<-c(rep(1,360),rep(2,360),rep(3,360),rep(4,360),rep(5,360),rep(6,360),rep(7,360),rep(8,360),rep(9,360),rep(10,360))

#Fix structure of data
str(CWR_1to10)

#Fix factorial variables
for (i in 1:7){
  CWR_1to10[,i]<-as.factor(CWR_1to10[,i])
}

#Fix continous variables
for (i in 9:13){
  CWR_1to10[,i]<-as.numeric(CWR_1to10[,i])
}

for (i in 15:21){
  CWR_1to10[,i]<-as.numeric(CWR_1to10[,i])
}

#Let's add a factorial term for generation
CWR_1to10$FacGen<-as.factor(CWR_1to10$Gen)

# Create an interaction term for treatment and media
CWR_1to10$Evol <- paste(CWR_1to10$Trt, CWR_1to10$Med, sep="")
CWR_1to10$Evol<- as.factor(CWR_1to10$Evol)

# Generate summary statistics of interest

# We will use growth rate (increase in surface area) as a proxy for fitness
CWR_1to10$Grt<-CWR_1to10$Pix1 - CWR_1to10$Pix0

# Calculate aggregation, a measure of how closely packed in fronds are. Lots of phenotypic variation in this trait, though its biological significance isn't clear....
# Will be calculated as surface area (# pixels) divided by perimeter, so that less aggregated wells (e.g. high perimeter) have a lower score
CWR_1to10$Agg<-CWR_1to10$Pix1/CWR_1to10$Per1

# We are also going to calculate greeness of the fronds. This could be a fitness-measure, or a trait plausibly associated to variation in growth rate/health.
CWR_1to10$Hlth<-CWR_1to10$Grn1/CWR_1to10$Pix1

# We will calculate mean frond size, just to see if this phenotype is affected by anything. Again, the biological significance of this is still unclear.
CWR_1to10$FrAve<-CWR_1to10$Pix1/CWR_1to10$Fr1

# Let's also calculate the change in number of fronds as an alternative measure of growth.
CWR_1to10$Grt_frd<-CWR_1to10$Fr1-CWR_1to10$Fr0

#Let's create a dataset containing only wells with plants. 
CWR_plt<-subset(CWR_1to10, CWR_1to10$Trt!= "Bac")
CWR_plt<-droplevels(CWR_plt)
```

SECTION 4: data assessment
Check normality of data.
```{r data assessment}
hist(CWR_1to10$Grt, breaks=20)
# That looks OK, let's try breaking this into populations

ggplot(data=CWR_1to10, aes(x = Grt)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ ., scales = "free")
# All have long tails to the right, and R especially has an overabundance of small values.

#Wwhat does this look like across generations?
ggplot(data=CWR_1to10, aes(x = Grt)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Gen, scales = "free")

#How about treatment?
ggplot(data=CWR_1to10, aes(x = Grt)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
# That looks the best, though I am still curious about logging the data.

CWR_1to10$Log_Grt<-log(CWR_1to10$Grt)
hist(CWR_1to10$Log_Grt, breaks=20)
#That looks better

ggplot(data=CWR_1to10, aes(x = Log_Grt)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ ., scales = "free")

ggplot(data=CWR_1to10, aes(x = Log_Grt)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
# OK yes, this now looks quite normal

## What about changes in frond number?
ggplot(data=CWR_1to10, aes(x = Grt_frd)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
#Looks OK

###########################OK, onto aggregation ###############

hist(CWR_1to10$Agg, breaks=20)

ggplot(data=CWR_1to10, aes(x = Agg)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
# That actually looks pretty OK.

###########################Health and Greenness? ###############

hist(CWR_1to10$Hlth, breaks=20)

ggplot(data=CWR_1to10, aes(x = Hlth)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
# That actually looks pretty OK.

ggplot(data=CWR_1to10, aes(x = Grn1)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
# Could log it?

ggplot(data=CWR_1to10, aes(x = log(Grn1))) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
# Looks better

###########################Frond size? ###############

hist(CWR_1to10$FrAve, breaks=20)

ggplot(data=CWR_1to10, aes(x = FrAve)) + geom_histogram(fill = "white", colour = "black") +  facet_grid(Pop ~ Trt ~ Med, scales = "free")
# Yeah that looks great.

#OK, so really growth is the only issue here. I've taken the log of it, and we will also try fitting a model with growth as a polynomial term

```

SECTION 5: modelling
OK, time for some models.
```{r models}

CWR_NA<-subset(CWR_plt, CWR_plt$Grt!= "NA")

# 1. Gen - continuous. Pos - edge, random effect. Grt - final pixel#, with initial pix as predictor. med*gen*trt. Pop - random
Mod_grt<-lmer(Pix1 ~ Pix0 + Med*Trt*Gen + (1|Pop) + (1|Edge), data= CWR_plt)

Mod_grt
Anova(Mod_grt, type=3)
# Significant effects of Pix0, Med, Gen, and Med:Gen.

# Test random effects one at a time? I'm not sure how to test both at the same time...
Mod_grt_pop<-lmer(Pix1 ~ Pix0 + Med*Trt*Gen + (1|Pop), data= CWR_plt)
exactRLRT(Mod_grt_pop)

Mod_grt_edge<-lmer(Pix1 ~ Pix0 + Med*Trt*Gen + (1|Edge), data= CWR_plt)
exactRLRT(Mod_grt_edge)

```

OK, so now we are going to generate Log Response Ratios for some traits...
Reference for calculating LRR's: Hedges, L. V., Gurevitch, J., & Curtis, P. S. (1999). The meta-analysis of response ratios in experimental ecology. Ecology, 80(4), 1150-1156.
```{r}
# Effect of inoculating plants with microbial communities
# Start by calculating statistics for each generation for each treatment.
melt_LRR<-melt(CWR_1to10, id.vars=c("Pop", "Gen","Trt","Med"), measure.vars= "Grt", na.rm = T)
Sum_LRR<- ddply(melt_LRR, c("Pop", "Med","Trt","Gen","variable"), summarise,
                mean = mean(value), sd = sd(value), count=length(value),
                sem = sd(value)/sqrt(length(value)))
Sum_LRR_plt<-subset(Sum_LRR, Sum_LRR$Trt!= "BAC")
Sum_LRR_plt<-droplevels(Sum_LRR_plt)
#LRR for Ctl media plants
pop<-3
med<-2
gen<-10

LRR_ctl<-data.frame(matrix(ncol=5, nrow=pop*med*gen))
names(LRR_ctl)<-c('Pop','Med','Gen','LRR','SE')
LRR_ctl$Pop<-c(rep('C', 20), rep('R', 20), rep('W', 20))
LRR_ctl$Med<-c(rep('Ctl', 10), rep('Salt', 10),rep('Ctl', 10), rep('Salt', 10),rep('Ctl', 10), rep('Salt', 10))
Gens<-c(1:10)
LRR_ctl$Gen<-c(rep(Gens, 6))

for (i in 1:10){
  #C, Ctl
  LRR_ctl[i,4]<-log(Sum_LRR_plt[i,6]/Sum_LRR_plt[i+10,6])
  LRR_ctl[i,5]<-sqrt(Sum_LRR_plt[i,9]/Sum_LRR_plt[i,6] + Sum_LRR_plt[i+10,9]/Sum_LRR_plt[i+10,6])
  #C, Salt
  LRR_ctl[i+10,4]<-log(Sum_LRR_plt[i+20,6]/Sum_LRR_plt[i+30,6])
  LRR_ctl[i+10,5]<-sqrt(Sum_LRR_plt[i+20,9]/Sum_LRR_plt[i+20,6] + Sum_LRR_plt[i+30,9]/Sum_LRR_plt[i+30,6])
  #R, Ctl
  LRR_ctl[i+20,4]<-log(Sum_LRR_plt[i+40,6]/Sum_LRR_plt[i+50,6])
  LRR_ctl[i+20,5]<-sqrt(Sum_LRR_plt[i+40,9]/Sum_LRR_plt[i+40,6] + Sum_LRR_plt[i+50,9]/Sum_LRR_plt[i+50,6])
  #R, Salt
  LRR_ctl[i+30,4]<-log(Sum_LRR_plt[i+60,6]/Sum_LRR_plt[i+70,6])
  LRR_ctl[i+30,5]<-sqrt(Sum_LRR_plt[i+60,9]/Sum_LRR_plt[i+60,6] + Sum_LRR_plt[i+70,9]/Sum_LRR_plt[i+70,6])  
  #W, Ctl
  LRR_ctl[i+40,4]<-log(Sum_LRR_plt[i+80,6]/Sum_LRR_plt[i+90,6])
  LRR_ctl[i+40,5]<-sqrt(Sum_LRR_plt[i+80,9]/Sum_LRR_plt[i+80,6] + Sum_LRR_plt[i+90,9]/Sum_LRR_plt[i+90,6])
  #W, Salt
  LRR_ctl[i+50,4]<-log(Sum_LRR_plt[i+100,6]/Sum_LRR_plt[i+110,6])
  LRR_ctl[i+50,5]<-sqrt(Sum_LRR_plt[i+100,9]/Sum_LRR_plt[i+100,6] + Sum_LRR_plt[i+110,9]/Sum_LRR_plt[i+110,6])
  
}
```